### 2023-01-18 Poisot_2023_guidelines_prediction_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@08949b36220530a13348d4076144bc29f375baf9 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{gensymb}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Guidelines for the prediction of species
interactions through binary classification}
\vskip 5em

%
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill

\begin{enumerate}
    \itemThe prediction of species interactions is gaining momentum as a
way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \itemUsing simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the ways to assemble
the training dataset.%
    \itemNeither classifier accuracy nor the area under the receiver
operating characteristic curve (ROC-AUC) are informative measures for
the performance of interaction prediction. The area under the
precision-recall curve (PR-AUC) is a fairer assessment of performance.
In some cases, even standard measures can lead to selecting a more
biased classifier because the effect of connectance is strong. The
amount of correction to apply to the training dataset depends on network
connectance, on the measure to be optimized, and only weakly on the
classifier.%
    \itemThese results reveal that training machines to predict networks
is a challenging task, and that in virtually all cases, the composition
of the training set needs to be fine-tuned before performing the actual
training. We discuss these consequences in the context of the low volume
of data.%
\end{enumerate}

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

Species interactions, forming ecological networks, are a backbone for
key ecological and evolutionary processes; yet enumerating all of the
interactions between \(S\) species is a daunting task, as it scales with
\(S^2\), \emph{i.e.} the squared species richness (Martinez, 1992).
Recent contributions to the field of ecological network prediction
(Becker et al., 2022; Pichler et al., 2020; Strydom et al., 2021)
highlight that although interactions can be predicted by adding
ecologically relevant information (in the form of, \emph{e.g.} traits),
we do not have robust guidelines as to how the predictive ability of
models recommending species interactions should be evaluated, nor about
how these models should be trained. Here, by relying on simple
derivations and a series of simulations, we formulate a number of such
guidelines, specifically for the case of binary classifiers derived from
thresholded values. Specifically, we conduct an investigation of the
models in terms of their skill (ability to make the right prediction),
bias (trends towards systematically over-predicting one class), class
imbalance (the relative number of cases representing interactions), and
show how these effects interact. We conclude on the fact that models
with the best interaction-scale predictive score do not necessarily
result in the most accurate representation of the true network.

The prediction of ecological interactions shares conceptual and
methodological issues with two fields in biology: species distribution
modelling (SDMs), and genomics. SDMs suffers from issues affecting
interactions prediction, namely low prevalence (due to sparsity of
observations/interactions) and data aggregation (due to bias in sampling
some locations/species). An important challenge lies in the fact that
the best measure to quantify the performance of a model is not
necessarilly a point of consensus (these methods, their interpretation,
and the way they are measured, are covered in depth in the next
section). In previous work, Allouche et al. (2006) suggested that
Cohen's \(\kappa\) agreement score (\(\kappa\) thereafter) was a better
test of model performance than the True Skill Statistic (TSS; which we
refer to as Youden's informedness thereafter); these conclusions were
later criticized by Somodi et al. (2017), who emphasized that
informedness is affected both by prevalence and bias. Although this work
offers recommendations about the comparison of models, it doesn't
establishes baselines or good practices for training on imbalanced
ecological data, or ways to remedy the imbalance. Steen et al. (2021)
show that, when applying spatial thinning (artificially re-balancing
observation data in space to avoid artifacts due to auto-correlation),
the best approach to train ML-based SDMs varies according to the
balancing of the dataset, and the evaluation measures used; there is no
single ``recipe'' that is guaranteed to give the best model. By contrast
to networks, SDMs have the advantage of being able to both thin datasets
to remove some of the sampling bias (\emph{e.g.} Inman et al., 2021),
but also to create pseudo-absences to inflate the number of supposed
negatives in the dataset (\emph{e.g.} Iturbide et al., 2015). These
powerful ways to remove data bias often have no analogue in networks,
removing one potential tool from our methodological toolkit, and making
the task of network prediction through classification potentially more
demanding, and more prone to underlying data biases.

An immense body of research on machine learning application to life
sciences is focused on genomics (which has very specific challenges, see
a recent discussion by Whalen et al., 2021); this sub-field has
generated recommendations that do not necessarily match the current
best-practices for SDMs, and therefore hint at the importance of
domain-specific guidelines. Chicco \& Jurman (2020) suggest using
Matthews correlation coefficient (MCC) over \(F_1\), as a protection
against over-inflation of predicted results; Delgado \& Tibau (2019)
advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
as the relative nature of \(\kappa\) means that a worse classifier can
be picked over a better one; similarly, Boughorbel et al. (2017)
recommend MCC over other measures of performance for imbalanced data, as
it has more desirable statistical properties. More recently, Chicco et
al. (2021) temper the apparent supremacy of the MCC, by suggesting it
should be replaced by Youden's informedness (also known as \(J\),
bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
in the dataset may not be representative of the actual imbalance. In a
way, the measures themselves need not be a strong focus for network
prediction, as they are routinely used in other field; the
discipline-specific question we seek to address is: `which metric should
be employed when predicting networks, and how to optimize it?'.

Species interaction networks are often under-sampled (Jordano, 2016a,
2016b), and this under-sampling is structured taxonomically (Beauchesne
et al., 2016), structurally (de Aguiar et al., 2019) and spatially
(Poisot, Bergeron, et al., 2021; Wood et al., 2015). As a consequence,
networks suffer from data deficiencies both within and between datasets.
This implies that the comparison of classifiers across space, when
undersampling varies locally (see \emph{e.g.} McLeod et al., 2021) is
non-trivial. Furthermore, the baseline value of classifiers performance
measures under various conditions of skill, bias, and prevalence, has to
be identified to allow researchers to evaluate whether their interaction
prediction model is indeed learning. Taken together, these
considerations highlight three specific issues for ecological networks.
First, what values of performance measures are indicative of a
classifier with no skill? This is particularly important as it can
evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training} or
\emph{testing}, and can we remedy it? Finally, because the low amount of
data on interaction makes a lot of imbalance correction methods (see
\emph{e.g.} Branco et al., 2015) hard to apply, which measures of model
performance can be optimized by sacrificing least amount of positive
interaction data?

A preliminary question is to examin the baseline performance of these
measures, \emph{i.e.} the values they would take on hypothetical
networks based on a classifier that has no-skill. It may sound
counter-intuitive to care so deeply about how good a classifier with
no-skill is, as by definition, is has no skill. The necessity of this
exercise has its roots in the paradox of accuracy: when the desired
class (``two species interact'') is rare, a model that gets less
ecologically performant by only predicting the opposite class (``these
two species do not interact'') sees its accuracy increase; because most
of the guesses have ``these two species do not interact'' as a correct
answer, a model that never predicts interactions would be right an
overwhelming majority of the time; it would also be utterly useless.
Herein lies the core challenge of predicting species interactions: the
extreme imbalance between classes makes the training of predictive
models difficult, and their validation even more so as we do not
reliably know which negatives are true. The connectance (the proportion
of realized interactions, usually the number of interactions divided by
the number of species pairs) of empirical networks is usually well under
20\%, with larger networks having a lower connectance (MacDonald et al.,
2020), and therefore being increasingly difficult to predict.

\hypertarget{a-primer-on-binary-classifier-evaluation}{%
\section{A primer on binary classifier
evaluation}\label{a-primer-on-binary-classifier-evaluation}}

Binary classifiers, which it to say, machine learning algorithms whose
answer is a binary value, are usually assessed by measuring properties
of their confusion matrix, \emph{i.e.} the contingency table reporting
true/false positive/negative hits. A confusion matrix is laid out as

\[\begin{pmatrix}
    \text{tp} & \text{fp} \\
    \text{fn} & \text{tn}
\end{pmatrix} \,.\]

In this matrix, tp is the number of times the model predicts an
interaction that exists in the network (true positive), fp is the number
of times the model predicts an interaction that does not exist in the
network (false positive), fn is the number of times the model fails to
predict an interaction that actually exists in the network (false
negatives), and tn is the number of times the model correctly predicts
that an interaction does not exist (true negatives). From these values,
we can derive a number of measures of model performance (see Strydom et
al., 2021 for a review of their interpretation in the context of
networks). At a coarse scale, a classifier is \emph{accurate} when the
trace of the matrix divided by the sum of the matrix is close to 1, with
other measures informing us on how the predictions fail.

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, and can therefore return a
value considered to be a pseudo-probability); in this case, the optimal
value below which predictions are assumed to be negative (\emph{i.e.}
the interaction does not exist) can be determined by picking a threshold
maximizing some value on the ROC or the PR curve. The area under these
curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
goodness of the classifier, and the ideal threshold is the point on
these curves that minimizes the tradeoff represented in these curves.
Saito \& Rehmsmeier (2015) established that the ROC-AUC is biased
towards over-estimating performance for imbalanced data; on the
contrary, the PR-AUC is able to identify classifiers that are less able
to detect positive interactions correctly, with the additional advantage
of having a baseline value equal to prevalence. Therefore, it is
important to assess whether these two measures return different results
when applied to ecological network prediction. The ROC curve is defined
by the false positive rate on the \(x\) axis, and the true positive rate
on the \(y\) axis, and the PR curve is defined by the true positive rate
on the \(x\) axis, and the positive predictive value on the \(y\) axis.

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013). The
choice of metrics with relevance to class-imbalanced problems is
fundamental, because as Japkowicz (2013) unambiguously concluded,
``relatively robust procedures used for unskewed data can break down
miserably when the data is skewed''. Following Japkowicz (2013), we
focus on two ranking metrics (the areas under the Receiver Operating
Characteristic and Precision Recall curves), and three threshold metrics
(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
show early on that it has undesirable properties).

The \(\kappa\) measure (Landis \& Koch, 1977) establishes the extent to
which two observers (the network and the prediction) agree, and is
measured as

\[
2\frac{tp\times tn - fn\times fp}{(tp+fp)\times (fp+tn)+(tn+fp)\times (tn+fn)} \,.
\]

Informedness (Youden, 1950) (also known as bookmaker informedness or the
True Skill Statistic) is \(\text{TPR}+\text{TNR}-1\), where
\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\). Informedness
can be used to find the optimal cutpoint in thresholding analyses
(Schisterman et al., 2005); indeed, the maximal informedness corresponds
to the point on the ROC curve that is closest to the perfect classifier
point. The formula for informedness is

\[\frac{tp}{tp+fn}+\frac{tn}{tn+fp}-1\,.\]

The MCC is defined as

\[
\frac{tp\times tn - fn\times fp}{\sqrt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
\]

Finally, \(F_1\) is the harmonic mean of precision (the chance that
interaction was correctly detected as such) and sensitivity (the ability
to correctly classify interactions), and is defined as

\[
2\frac{tp}{2\times tp + fp + fn}\,.
\]

One noteworthy fact is that \(F_1\) and MCC have ties to the PR curve
(being close to the expected PR-AUC), and that informedness has ties to
the ROC curve (whereby the threshold maximizing informedness is also the
point of maximal inflection on the ROC curve). One important difference
between ROC and PR is that the later does not prominently account for
the size of the true negative compartments: in short, it is more
sensitive to the correct positive predictions. In a context of strong
imbalance, PR-AUC is therefore a more stringent test of model
performance.

\hypertarget{baseline-values-for-the-threshold-metrics}{%
\section{Baseline values for the threshold
metrics}\label{baseline-values-for-the-threshold-metrics}}

In this section, we will assume a network with connectance equal to a
scalar \(\rho\), \emph{i.e.} having \(\rho S^2\) interactions (where
\(S\) is the species richness), and \((1-\rho) S^2\) non-interactions.
Therefore, the vector describing the \emph{true} state of the network
(assumed to be an unweighted, directed network) is a column vector
\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

\hypertarget{confusion-matrix-with-skill-and-bias}{%
\subsection{Confusion matrix with skill and
bias}\label{confusion-matrix-with-skill-and-bias}}

In order to write the values of the confusion matrix for a hypothetical
classifier, we need to define two characteristics: its skill, and its
bias. Skill, here, refers to the propensity of the classifier to get the
correct answer (\emph{i.e.} to assign interactions where they are, and
to not assign them where they are not). A no-skill classifier guesses at
random, \emph{i.e.} it will guess interactions with a probability
\(\rho\). The predictions of a no-skill classifier can be expressed as a
row vector \(\mathbf{p}^T = [\rho, (1-\rho)]\). The confusion matrix
\(\mathbf{M}\) for a no-skill classifier is given by the element-wise
(Hadamard, outer) product of these vectors
\(\mathbf{o} \odot \mathbf{p}\), \emph{i.e.}

\[
\mathbf{M} = \begin{pmatrix}
    \rho^2 & \rho (1-\rho) \\
    (1-\rho) \rho & (1-\rho)^2
\end{pmatrix} \,.
\]

In order to regulate the skill of this classifier, we can define a skill
matrix \(\mathbf{S}\) with diagonal elements equal to \(s\), and
off-diagonal elements equal to \((1-s)\), which allows to regulate how
many predictions are wrong, under the assumption that the bias is the
same (\emph{i.e.} the classifier is as likely to make a false positive
or a false negative). The skill-adjusted confusion matrix is
\(\mathbf{M} \odot \mathbf{S}\), \emph{i.e.}

\[
\begin{pmatrix}
    \rho^2 & \rho (1-\rho) \\
    (1-\rho) \rho & (1-\rho)^2
\end{pmatrix} \odot \begin{pmatrix}
    s & (1-s) \\
    (1-s) & s
\end{pmatrix} \,.
\]

When \(s=0\), \(\text{Tr}(\mathbf{M}) = 0\) (the classifier is
\emph{always} wrong), when \(s=0.5\), the classifier is no-skill and
guesses at random, and when \(s=1\), the classifier is perfect.

The second element we can adjust in this hypothetical classifier is its
bias, specifically its tendency to over-predict interactions. Like
above, we can do so by defining a bias matrix \(\mathbf{B}\), where
interactions are over-predicted with probability \(b\), and express the
final classifier confusion matrix as
\(\mathbf{M}\odot \mathbf{S}\odot \mathbf{B}\), \emph{i.e.}

\[
\begin{pmatrix}
    \rho^2 & \rho (1-\rho) \\
    (1-\rho) \rho & (1-\rho)^2
\end{pmatrix} \odot \begin{pmatrix}
    s & (1-s) \\
    (1-s) & s
\end{pmatrix} \odot \begin{pmatrix}
    b & b \\
    (1-b) & (1-b)
\end{pmatrix}\,.
\]

The final expression for the confusion matrix in which we can regulate
the skill and the bias is

\[
\mathbf{C} = \begin{pmatrix}
    s\times b\times \rho^2 & (1-s)\times b\times \rho (1-\rho) \\
    (1-s)\times (1-b)\times (1-\rho) \rho & s\times (1-b)\times (1-\rho)^2
\end{pmatrix} \,.
\]

In all further simulations, the confusion matrix \(\mathbf{C}\) is
transformed so that it sums to unity, \emph{i.e.} the entries are the
\emph{proportions} of guesses.

\hypertarget{what-are-the-baseline-values-of-performance-measures}{%
\subsection{What are the baseline values of performance
measures?}\label{what-are-the-baseline-values-of-performance-measures}}

In this section, we will change the values of \(b\), \(s\), and
\(\rho\), and report how the main measures discussed in the introduction
(MCC, \(F_1\), \(\kappa\), and informedness) respond. Before we do so,
it is important to explain why we will not focus on accuracy too much.
Accuracy is the number of correct predictions
(\(\text{Tr}(\mathbf{C})\)) divided by the sum of the confusion matrix.
For a no-skill, no-bias classifier, accuracy is equal to
\(\rho^2 + (1-\rho)^2\); for \(\rho = 0.05\), this is \(\approx 0.90\),
and for \(\rho = 0.01\), this is equal to \(\approx 0.98\). In other
words, the values of accuracy are high enough to be uninformative (for
\(\rho\) small, \(\rho^2 \ll (1-\rho)^2\)). More concerning is the fact
that introducing bias changes the response of accuracy in unexpected
ways. Assuming a no-skill classifier, the numerator of accuracy becomes
\(b\rho^2 + (1-b)(1-\rho)^2\), which increases when \(b\) is low, which
specifically means that at equal skill, a classifier that under-predicts
interactions will have higher accuracy than an un-biased classifier
(because the value of accuracy is dominated by the size of tn, which
will increase). These issues are absent from balanced accuracy, but
should nevertheless lead us to not report accuracy as the primary
measure of network prediction success; moving forward, we will focus on
other measures.

In order to examine how MCC, \(F_1\), \(\kappa\), and informedness
change w.r.t. the imbalance, skill, and bias, we performed a grid
exploration of the values of \(\text{logit}(s)\) and \(\text{logit}(b)\)
linearly from \(-10\) to \(10\); \(\text{logit}(x) = -10\) means that
\(x\) is essentially 0, and \(\text{logit}(x) = 10\) means it is
essentially 1 -- this choice was motivated by the fact that most
responses are non-linear with regards to bias and skill. The values or
\(\rho\) were taken linearly in \(]0, 0.5]\), which is within the range
of connectance for species interaction networks. Note that at this
point, there is no network model to speak of; the confusion matrix we
discuss can be obtained for any classification task. Based on the
previous discussion, the desirable properties for a measure of
classifier success should be: an increase with classifier skill,
especially at low bias; a hump-shaped response to bias, especially at
high skill, and ideally centered around \(\text{logit}(b)=0\); an
increase with prevalence up until equiprevalence is reached.

\begin{figure}
\hypertarget{fig:bias}{%
\centering
\includegraphics{figures/changing-bias.png}
\caption{Consequences of changing the classifier skills (\(s\)) and bias
(\(s\)) for a connectance \(\rho=0.15\), on \(F_1\), informedness, MCC,
and \(\kappa\). Accuracy increases with skill, but also increases when
the bias tends towards estimating \emph{fewer} interactions (this
follows from the derivations in the text, not shown in the figure).
Interestingly, \(\kappa\) responds as expected to skill (being negative
whenever \(s < 0.5\)), and peaks for values of \(b \approx 0.5\);
nevertheless, the value of bias for which \(\kappa\) is maximized in
\emph{not} \(b=0.5\), but instead increases with classifier skill. In
other words, at equal skill, maximizing \(\kappa\) would lead to select
a \emph{more} biased classifier.}\label{fig:bias}
}
\end{figure}

In fig.~\ref{fig:bias}, we show that none of the four measures satisfy
all the considerations at once: \(F_1\) increases with skill, and
increases monotonously with bias; this is because \(F_1\) does not
account for true negatives, and the increase in positive detection masks
the over-prediction of interactions. Informedness varies with skill,
reaching 0 for a no-skill classifier, but is entirely unsensitive to
bias. Both MCC and \(\kappa\) have the same behavior, whereby they
increase with skill. \(\kappa\) peaks at increasing values of bias for
increasing skill, \emph{i.e.} is likely to lead to the selection of a
classifier that over-predicts interactions. By contract, MCC peaks at
the same value, regardless of skill, but this value is not
\(\text{logit}(b)=0\): unless at very high classifier skill, MCC risks
leading to a model that over-predicts interactions. In
fig.~\ref{fig:connectance}, we show that all measures except \(F_1\)
give a value of 0 for a no-skill classifier, and are forced towars their
correct maximal value when skill changes (\emph{i.e.} a more connected
networks will have higher values for a skilled classifierd, and lower
values for a classifier making mostly mistakes).

\begin{figure}
\hypertarget{fig:connectance}{%
\centering
\includegraphics{figures/changing-connectance.png}
\caption{As in fig.~\ref{fig:bias}, consequences of changing connectance
for different levels of classifier skill, assuming no classifier bias.
Informedness, \(\kappa\), and MCC do increase with connectance, but only
when the classifier is not no-skill; by way of contrast, a more
connected network will give a higher \(F_1\) value even with a no-skill
classifier.}\label{fig:connectance}
}
\end{figure}

These two analyses point to the following recommendations: MCC is indeed
more appropriate than \(\kappa\), as although sensitive to bias, it is
sensitive in a consistent way. Informedness is appropriate at
discriminating between different skills, but confounded by bias. As both
of these measures bring valuable information on the model behavior, we
will retain them for future analyses. \(F_1\) is increasing with bias,
and should not be prioritized to evalue the performance of the model.
The discussion of sensitivity to bias should come with a domain-specific
caveat: although it is likely that interactions documented in ecological
networks are correct, a lot of non-interactions are simply unobserved;
as predictive models are used for data-inflation (\emph{i.e.} the
prediction of new interactions), it is not necessarily a bad thing in
practice to select models that predict more interactions than the
original dataset, because the original dataset misses some interactions.
Furthermore, the weight of positive interactions could be adjusted if
some information about the extent of undersampling exists (\emph{e.g.}
Branco et al., 2015). In a recent large-scale imputation of interactions
in the mammal-virus networks, Poisot, Ouellet, et al. (2021) for example
estimated that 93\% of interactions are yet to be documented.

\hypertarget{numerical-experiments-on-training-strategy}{%
\section{Numerical experiments on training
strategy}\label{numerical-experiments-on-training-strategy}}

In the following section, we will generate random bipartite networks,
and train four binary classifiers (as well as an ensemble model using
the sum of ranged outputs from the component models) on 50\% of the
interaction data. In practice, testing usually uses 70\% of the total
data; for ecological networks, where interactions are sparse \emph{and}
the number of species is low, this may not be the best solution, as the
testing set becomes constrained not by the \emph{proportion} of
interactions, but by their \emph{number}. Preliminary experiments using
different splits revealed no qualitative change in the results. Networks
are generated by picking a random infectiousness trait \(v_i\) for 100
species (from a beta distribution \(B(\alpha=6,\beta=8)\) distribution),
and a resistance trait \(h_j\) for 100 species (from
\(B(\alpha=2,\beta=8)\) distribution). There is an interaction between
\(i\) and \(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\)
is a constant regulating the connectance of the network (visual
exploration of the parameters show that there is an almost 1:1
relationship between \(\xi\) and connectance), and varies uniformly in
\([0.05, 0.35]\). This model gives fully interval networks that are
close analogues to the bacteria--phage model of Weitz et al. (2005),
with both a modular structure and a non-uniform degree distribution.
This dataset is easy for almost any algorithm to learn: when trained
with features \([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the
interactions between \(i\) and \(j\), all four models presented below
were able to reach almost perfect predictions all the time (data not
presented here) -- this is in part because the rule (there is maximum
value of the distance between traits for which there is an interaction)
is fixed for all interactions, and any method able to learn non-linear
relationships should infer it without issues. In order to make the
problem more difficult to solve, we use \([v_i, h_j]\) as a feature
vector (\emph{i.e.} the traits on which the models are trained), and
therefore the models will have to uncover that the rule for interaction
is \(\text{abs}(v_i, h_j) \le \xi\). The models therefore all have the
following form, where \(i_{i,j}\) is an interaction from species \(i\)
to species \(j\):

\[
\begin{bmatrix}
    i_{1,1} \\
    i_{1,2} \\
    \vdots \\
    i_{m,n-1} \\
    i_{m,n}
\end{bmatrix}
\propto
\begin{bmatrix}
           v_{1} & h_{1} \\
           v_{1} & h_{2} \\
           \vdots & \vdots \\
           v_{m} & h_{n-1} \\
           v_{m} & h_{n}
         \end{bmatrix}
\]

The training sample is composed of a random pick of up to 50\% of the
\(10^4\) possible entries in the network, \emph{i.e.} \(n=5000\). Out of
these interactions, we pick a proportion \(\nu\) (the training set
balance) to be positive, so that the training set has \(\nu n\)
interactions, and \((1-\nu) n\) non-interactions. We vary \(\nu\)
uniformly in \(]0,1[\). This allows to evaluate how the measures of
binary classification performance respond to artificially rebalanced
dataset for a given network connectance. The rest of the dataset is used
as a testing set, on which all further measures are calculated. Note
that although the training set is balanced arbitrarily, the testing set
is assembled so that it has the exact connectance of the entire network;
this ensures that the model is evaluated under the class imbalance where
the predictions will be made, which represents a more meaningful
evaluation. Furthermore, to avoid artifacts due to different sizes of
the training and testing set within a single network, the number of
entries in both sets are equal. Note also that although the simulated
networks are bipartite, the algorithms have no ``knowledge'' of the
network structure, and simply look at pairs of species; therefore, the
approach outlined here would also work for unipartite networks.

The dataset used for numerical experiments is composed of a grid of 35
values of connectance (from 0.011 to 0.5) and 35 values of \(\nu\) (from
0.02 to 0.98); for each pair of values, 500 networks are generated and
predicted. For each network, we train four machines: a trait-based k-NN
(\emph{e.g.} Desjardins-Proulx et al., 2017), a regression tree, a
regression random forest, and a boosted regression tree; the later three
methods are turned into classifiers using thresholding, which oftentimes
provides better results than classification when faced with class
imbalance (Hong et al., 2016). Following results from Pichler et al.
(2020), linear models have not been considered (in any way, the
relationship in the simulated networks is non-linear). The point of
these numerical experiments is \emph{not} to recommend the best model
(this is likely problem-specific), but to highlight a series of
recommendations that would work for supervised learning tasks. All
models were taken from the \texttt{MLJ.jl} package (Blaom et al., 2020;
Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et al., 2017). All
machines use the default parameterization; this is an obvious deviation
from best practices, as the hyperparameters of any machine require
training before its application on a real dataset. As we use 612500 such
datasets, this would require over 2 millions unique instances of
tweaking the hyperparameters, which is prohibitive from a computing time
point of view. An important thing to keep in mind is that the problem we
simulate has been designed to be simple to solve: we expect all machines
with sensible default parameters to fare well --- the results presented
in the later sections show that this assumption is warranted, and we
further checked that the models do not overfit by ensuring that there is
never more than 5\% of difference between the accuracy on the training
and testing sets. All machines return a quantitative prediction, usually
(but not necessarily) in \([0,1]\), which is proportional (but not
necessarily linearly) to the probability of an interaction between \(i\)
and \(j\). The ROC-AUC and PR-AUC (and therefore the thresholds) can be
measured by integrating over the domain of the values return by each
machine, but in order to make the average-based ensemble model more
meaningful, all predictions are expressed in \([0,1]\).

In order to pick the best confusion matrix for a given trained machine,
we performed a thresholding approach using 500 steps on predictions from
the testing set, and picking the threshold that maximized Youden's
informedness. During the thresholding step, we measured the area under
the receiver operating characteristic (ROC-AUC) and precision-recall
(PR-AUC) curves, as measures of overall performance over the range of
returned values. We report the ROC-AUC and PR-AUC, as well as a suite of
other measures as introduced in the next section, for the best
threshold. The ensemble model was generated by summing the predictions
of all component models on the testing set (ranged in \([0,1]\)), then
put through the same thresholding process. The complete code to run the
simulations is available at \texttt{10.17605/OSF.IO/JKEWD}.

After the simulations were completed, we removed all runs (\emph{i.e.}
triples of model, \(\xi\), and \(\nu\)) for which at least one of the
following conditions was met: the accuracy was 0, the true positive or
true negative rates were 0, the connectance was larger than 0.25. This
removes both the obviously failed model runs, and the networks that are
more densely connected compared to the connectance of empirical food
webs (and are therefore less difficult to predict, being less
imbalanced; preliminary analyses of data with a connectance larger than
0.3 revealed that all machines reached consistently high performance).

\hypertarget{effect-of-training-set-balance-on-performance}{%
\subsection{Effect of training set balance on
performance}\label{effect-of-training-set-balance-on-performance}}

In fig.~\ref{fig:biasco}, we present the response of two thresholding
measures (PR-AUC and ROC-AUC) and two ranking measures (Informedness and
MCC) to a grid of 35 values of training set balance, and 35 values of
connectance, for the four component models as well as the ensemble.
ROC-AUC is always high, and does not vary with training set balance. On
the other hand, PR-AUC shows very strong responses, increasing with
training set balance. It is notable here that two classifiers that
seemed to be performing well (Decision Tree and Random Forest) based on
their MCC are not able to reach a high PR-AUC even at higher
connectances. All models reached a higher performance on more connected
networks, and using more balanced training sets. In all cases,
informedness was extremely high, which is an expected consequence of the
fact that this is the value we optimized to determine the cutoff. MCC
increased with training set balance, although this increase became less
steep with increasing connectance. Three of the models (kNN, decision
tree, and random forest) only increased their PR-AUC sharply when the
training set was heavily imbalanced towards more interactions.
Interestingly, the ensemble almost always outclassed its component
models. For larger connectances (less difficult networks to predict, as
they are more balanced), MCC and informedness stared decreasing when the
training set bias got too close to one, suggesting that a training set
balance of 0.5 may often be appropriate if these measures are the one to
optimize.

\begin{figure}
\hypertarget{fig:biasco}{%
\centering
\includegraphics{figures/bias_by_connectance.png}
\caption{Response of MCC, Informedness, ROC-AUC, and PR-AUC to changes
in the training set balance (on the \(x\) axis) for a series of
increasing connectances (color). All of these values approach 1 for a
good model, but should be lower when the prediction is more difficult.
Informedness is consistently high, and by contrast, MCC increases with
additional training set balance. Across all models, training on a more
connected network is easier. ROC-AUC is consistently high, and therefore
not properly able to separate good from poor classifiers. On the other
hand, PR-AUC responds to changes in the training set.}\label{fig:biasco}
}
\end{figure}

Based on the results presented in fig.~\ref{fig:biasco}, it seems that
informedness and ROC-AUC are not necessarily able to discriminate
between good and bad classifiers (although this result may be an
artifact for informedness, as it has been optimized when thresholding).
On the other hand, MCC and PR-AUC show a strong response to training set
balance, and may therefore be more useful at model comparison.

\hypertarget{required-amount-of-positives-to-get-the-best-performance}{%
\subsection{Required amount of positives to get the best
performance}\label{required-amount-of-positives-to-get-the-best-performance}}

The previous results revealed that the measure of classification
performance responds both to the bias in the training set \emph{and} to
the connectance of the network; from a practical point of view,
assembling a training set requires one to withhold positive information,
which in ecological networks are very scarce (and typically more
valuable than negatives, on which there is a doubt). For this reason,
across all values of connectance, we measured the training set balance
that maximized a series of performance measures. When this value is
high, the training set needs to skew more positive in order to get a
performant model; when this value is about 0.5, the training set needs
to be artificially balanced to optimize the model performance. These
results are presented in fig.~\ref{fig:optimbias}.

\begin{figure}
\hypertarget{fig:optimbias}{%
\centering
\includegraphics{figures/optimal_bias.png}
\caption{Value of the optimal training set balance for the different
models and measures evaluated here, over a range of connectances.
Informedness was reliably maximized for balanced training sets, and kept
this behavior across models. For other measures, larger connectances in
the true network allowed lower biases in the training set. In a large
number of cases, ``over-correcting'' by having training sets with more
than half instances representing interactions would maximize the values
of the model performance measures.}\label{fig:optimbias}
}
\end{figure}

The more ``optimistic'' measures (ROC-AUC and informedness) required a
biasing of the dataset from about 0.4 to 0.75 to be maximized, with the
amount of bias required decreasing only slightly with the connectance of
the original network. MCC and PR-AUC required values of training set
balance from 0.75 to almost 1 to be optimized, which is in line with the
results of the previous section, \emph{i.e.} they are more stringent
tests of model performance. These results suggest that learning from a
dataset with very low connectance can be a different task than for more
connected networks: it becomes increasingly important to capture the
mechanisms that make an interaction \emph{exist}, and therefore having a
slightly more biased training dataset might be beneficial. As
connectance increases, the need for biased training sets is less
prominent, as learning the rules for which interactions \emph{do not}
exist starts gaining importance.

\begin{figure}
\hypertarget{fig:optimvalue}{%
\centering
\includegraphics{figures/optimal_value.png}
\caption{When trained on their optimally biased training set, most
models were able to maximize their performance; this is not true when
measuring PR-AUC for decision tree, k-NN, and to a lower extent RF. The
ensemble had a consistently high performance despite incorporating
low-performing models.}\label{fig:optimvalue}
}
\end{figure}

When trained at their optimal training set balance, connectance still
had a significant impact on the performance of some machines
(fig.~\ref{fig:optimvalue}). Notably, Decision Tree, and k-NN, as well
as Random forest to a lower extent, had low values of PR-AUC. In all
cases, the Boosted Regression Tree was reaching very good predictions
(especially for connectances larger than 0.1), and the ensemble was
almost always scoring perfectly. This suggests that all the models are
biased in different ways, and that the averaging in the ensemble is able
to correct these biases. We do not expect this last result to have any
generality, and provide a discussion of a recent example in which the
ensemble was performing worse than its components models.

\hypertarget{do-better-classification-accuracy-result-in-more-realistic-networks}{%
\section{Do better classification accuracy result in more realistic
networks?}\label{do-better-classification-accuracy-result-in-more-realistic-networks}}

In this last section, we generate a network using the same model as
before, with \(S_1, S_2 = 50, 80\) species, a connectance of
\(\approx 0.16\) (\(\xi = 0.19\)), and a training set balance of
\(0.5\), as fig.~\ref{fig:optimbias} suggests this is the optimal
training set balance for this range of connectance. The prediction made
on the complete dataset is presented in fig.~\ref{fig:ecovalid}.

\begin{figure}
\hypertarget{fig:ecovalid}{%
\centering
\includegraphics{figures/valid_ensemble.png}
\caption{Visualisation of the raw (un-thresholded) models predictions
for one instance of a network prediction problem (shown in the
``Dataset'' panel). Increasing the value of the \(\xi\) parameter would
make the diagonal structure ``broader'', leading to more interactions. A
visual inspection of the results is important, as it highlights how some
models can ``miss'' parts of the network; by combining them in an
ensemble, these gaps compensate one another, and lead (in this case) to
a better prediction.}\label{fig:ecovalid}
}
\end{figure}

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness
{[}\(\eta\); Bastolla et al. (2009){]}, modularity {[}\(Q\); Barber
(2007){]}, asymmetry {[}\(A\); Delmas et al. (2018){]}, and Jaccard
network dissimilarity (Canard et al., 2014). This process was repeated
250 times, and the results are presented in tbl.~\ref{tbl:comparison}.
The k-NN model is an interesting instance here: it produces the network
that looks the most like the original dataset, despite having the lowest
PR-AUC, suggesting it hits high recall at the cost of low precision. The
ensemble was able to reach a very high PR-AUC (and a very high ROC-AUC),
which translated into more accurate reconstructions of the structure of
the network (with the exception of modulairty, which is underestimated
by \(0.03\)). This result bears elaborating. Measures of model
performance capture how much of the interactions and non-interactions
are correctly identified. As long as these predictions are not perfect,
some interactions will be predicted at the ``wrong'' position in the
network; these measures cannot describe the structural effect of these
mistakes. On the other hand, measures of network structure can have the
same value with interactions that fall at drastically different
positions; this is in part because a lot of these measures covary with
connectance, and in part because as long as these values are not 0 or
their respective maximum, there is a large number of network
configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013). In this specif example, the combination of individually
``adequate'' models resulted in an extremely strong ensemble, suggesting
that the correct prediction of interactions (as measured by MCC, Inf.,
ROC-AUC, and PR-AUC) \emph{and} network properties is indeed a feasible
task under appropriately hyper-parameterized models.

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1548}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0833}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
five network structure metrics, for 500 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MCC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Inf.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
PR-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Conn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(\eta\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(Q\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(A\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Jaccard
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MCC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Inf.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
PR-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Conn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(\eta\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(Q\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(A\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Jaccard
\end{minipage} \\
\midrule()
\endhead
Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
\textbf{0.42} & 0.1 \\
BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 & 0.19 \\
Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06} \\
k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
\textbf{0.61} & \textbf{0.39} & \textbf{0.42} & \textbf{0.06} \\
\emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
\textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
\textbf{0.06} \\
\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0 \\
\bottomrule()
\end{longtable}

\hypertarget{guidelines-for-the-assessment-of-network-predictive-models}{%
\section{Guidelines for the assessment of network predictive
models}\label{guidelines-for-the-assessment-of-network-predictive-models}}

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidentally correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble, we show that
informedness and ROC-AUC are consistently high on network data, whereas
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not necessarilly the models with the closest
reconstructed network structure. We discuss these results in the context
of establishing guidelines for the prediction of ecological
interactions.

It is noteworthy that the ensemble model was systematically better than
the component models. We do not expect that ensembles will \emph{always}
be better than single models. Networks with different structures than
the one we simulated here may respond in different ways, especially if
the rules are fuzzier than the simple rule we used here. In a recent
multi-model comparison involving supervised and unsupervised learning,
Becker et al. (2022) found that the ensemble was \emph{not} the best
model, and was specifically under-performing compared to models using
biological traits. This may be because the dataset of Becker et al.
(2022) was known to be under-sampled, and so the network alone contained
less information than the combination of the network and species traits.
There is no general conclusion to draw from either these results or
ours, besides reinforcing the need to be pragmatic about which models
should be included in the ensemble, and whether to use an ensemble at
all. In a sense, the surprising performance of the ensemble model should
form the basis of the first broad recommendation: optimal training set
balance and its interaction with connectance and the specific binary
classifier used is, in a sense, an hyperparameter that should be
assessed following the approach outlined in this manuscript. The
distribution of results in fig.~\ref{fig:optimbias} and
fig.~\ref{fig:optimvalue} show that there are variations around the
trend, and multiple models should probably be trained on their
``optimal'' training/testing set, as opposed to the same ones.

The results presented here highlight an interesting paradox: although
the k-NN model was ultimately able to get a correct estimate of network
structure (see tbl.~\ref{tbl:comparison} and fig.~\ref{fig:ecovalid}),
it ultimately remains a poor classifier, as evidenced by its low PR-AUC.
This suggests that the goal of predicting \emph{interactions} and
predicting \emph{networks} may not always be solvable in the same way --
of course a perfect classifier of interactions would make a perfect
network prediction; indeed, the best scoring predictor of interactions
(the ensemble model) had the best prediction of network structure. The
tasks of predicting networks structure and of predicting interactions
within networks are essentially two different ones. For some
applications (\emph{e.g.} comparison of network structure across
gradients), one may care more about a robust estimate of the structure,
at the cost at putting some interactions at the wrong place. For other
applications (\emph{e.g.} identifying pairs of interacting species), one
may conversely care more about getting as many pairs right, even though
the mistakes accumulate in the form of a slightly worse estimate of
network structure. How these two approaches can be reconciled is
something to evaluate on a case-by-case basis, especially since there is
no guarantee that an ensemble model will always be the most precise one.
Despite this apparent tension at the heart of the predictive exercise,
we can use the results presented here to suggest a number of guidelines.

First, because we have more trust in reported interactions than in
reported absences of interactions (which are overwhelmingly
\emph{pseudo}-absences), we can draw on previous literature to recommend
informedness as a measure to decide on a threshold for binary
classification (Chicco et al., 2021); this being said, because
informedness is insensitive to bias (although it is a linear measure of
skill), the overall model performance is better evaluated through the
use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
\(F_1\) is monotonously sensitive to classifier bias
(fig.~\ref{fig:bias}) and network connectance
(fig.~\ref{fig:connectance}), MCC should be prefered as a measure of
model evaluation and comparison. When dealing with multiple models, we
therefore suggest to find the optimal threshold using informedness, and
to pick the best model using MCC (assuming one does not want to use an
ensemble model).

Second, accuracy alone should not be the main measure of model
performance, but rather an expectation of how well the model should
behave given the class balance in the set on which predictions are made;
this is because, as derived earlier, the expected accuracy for a
no-skill no-bias classifier is \(\rho^2 + (1-\rho)^2\) (where \(\rho\)
is the class balance), which will most often be large. This pitfall is
notably illustrated in a recent food-web model (Caron et al., 2022)
wherein the authors, using a training set of \(n = 10^4\) with only 100
positive interactions (representing 0.1\% of the total interactions),
reached a good accuracy. Reporting a good accuracy is not informative,
especially when accuracy isn't (i) compared to the baseline expected
value under the given class balance, and (ii) interpreted in the context
of a measure that is not sensitive to the chance prediction of many
negatives (like MCC).

Third, because the PR-AUC responds more to network connectance
(fig.~\ref{fig:optimvalue}) and training set imbalance
(fig.~\ref{fig:optimbias}) than ROC-AUC, it should be used as a measure
of model performance over the ROC-AUC. This is not to say that ROC-AUC
should be discarded (in fact, a low ROC-AUC is undoubtedly a sign of an
issue with the model), but that its interpretation should be guided by
the PR-AUC value. Specifically, a high ROC-AUC is not informative, as it
can be associated to a low PR-AUC (see \emph{e.g.} Random Forest in
tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015). We therefore
expect to see high ROC-AUC values, and then to pick the model that
maximizes the PR-AUC value. Taken together with the previous two
guidelines, we strongly encourage to (i) ensure that accuracy and
ROC-AUC are high (in the case of accuracy, higher than expected under
no-skill no-bias situation), and (ii) to discuss the performance of the
model in terms of the most discriminant measures, \emph{i.e.} PR-AUC and
MCC.

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set, because
it is an ecologically relevant value. In the approach outlined here, we
treat the class imbalance of the training set as an hyper-parameter, but
\emph{test} the model on a set that has the same class imbalance as the
actual dataset. This is an important distinction, as it ensure that the
prediction environment matches the testing environment (as we cannot
manipulate the connectance of the empirical dataset on which the
predictions will be made), and so the values measured on the testing set
(or validation set if the data volume allows one to exists) can be
directly compared to the values for the actual prediction. A striking
result from fig.~\ref{fig:optimbias} is that Informedness was almost
always maximal at 50/50 balance (regardless of connectance), whereas MCC
required \emph{more} positives to be maximized when connectance
\emph{increases}, matching the idea that it is a more stringent measure
of performance. This has an important consequence in ecological
networks, for which the pool of positive cases (interactions) to draw
from is typically small: the most parsimonious measure (\emph{i.e.} the
one requiring to discard the least amount of interactions to train the
model) will give the best validation potential, and in this light is
very likely informedness {[}maximizing informedness is, in fact, the
generally accepted default for imbalanced classification regardless of
the problem domain; Schisterman et al. (2005){]}. This last result
further strengthens the assumption that the amount of bias \emph{is} an
hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to models with lower performance; for this reason, it makes sense
to not train all models on the same training/testing set, but rather to
optimize the set composition for each of them.

One key element for real-life data that can make the prediction exercise
more tractable is that some interactions can safely be assumed to be
impossible; indeed, a lot of networks can be reasonably well described
using a stochastic block model (\emph{e.g.} Xie et al., 2017). In
ecological networks, this can be due to spatial constraints (Valdovinos,
2019), or to the long-standing knowledge that some links are
``forbidden'' due to traits (Olesen et al., 2011) or abundances (Canard
et al., 2014). The matching rules (Olito \& Fox, 2015; Strona \& Veech,
2017) can be incorporated in the model either by adding compatibility
traits, or by \emph{only} training the model on pairs of species that
are not likely to be forbidden links. Knowledge of true negative
interactions could be propagated in training/testing sets that have true
negatives, and in this situation, it may be possible to use the more
usual 70/30 split for training/testing folds as the need to protect
against potential unbalance is lowered. Besides forbidden links, a
real-life case that may arise is multi-interaction or multi-layer
networks (Pilosof et al., 2017). These can be studied using the same
general approach outlined here, either by assuming that pairs of species
can interact in more than one way (wherein one would train a model for
each type of interaction, based on the relevant predictors), or by
assuming that pairs of species can only have one type of interaction
(wherein this becomes a multi-label classification problem).

\textbf{Acknowledgements:} We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. We thank Colin J. Carlson, Michael D. Catchen, Giulio Valentino
Dalla Riva, and Tanya Strydom for inputs on earlier versions of this
manuscript. This research was enabled in part by support provided by
Calcul Québec (www.calculquebec.ca) through the Narval general purpose
cluster. TP is supported by the Fondation Courtois, a NSERC Discovery
Grant and Discovery Acceleration Supplement, by funding to the Viral
Emergence Research Initiative (VERENA) consortium including NSF BII
2021909, and by a grant from the Institut de Valorisation des Données
(IVADO).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Allouche2006AssAcc}{}}%
Allouche, O., Tsoar, A., \& Kadmon, R. (2006). Assessing the accuracy of
species distribution models: Prevalence, kappa and the true skill
statistic (TSS). \emph{Journal of Applied Ecology}, \emph{43}(6),
1223--1232. \url{https://doi.org/10.1111/j.1365-2664.2006.01214.x}

\leavevmode\vadjust pre{\hypertarget{ref-Barber2007ModCom}{}}%
Barber, M. J. (2007). Modularity and community detection in bipartite
networks. \emph{Physical Review E}, \emph{76}(6), 066102.
\url{https://doi.org/10.1103/PhysRevE.76.066102}

\leavevmode\vadjust pre{\hypertarget{ref-Bastolla2009ArcMut}{}}%
Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
B., \& Bascompte, J. (2009). The architecture of mutualistic networks
minimizes competition and increases biodiversity. \emph{Nature},
\emph{458}(7241), 1018--1020. \url{https://doi.org/10.1038/nature07950}

\leavevmode\vadjust pre{\hypertarget{ref-Beauchesne2016ThiOut}{}}%
Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
(2016). Thinking Outside the Box--predicting Biotic Interactions in
Data-poor Environments. \emph{Vie Et Milieu-Life and enVironment},
\emph{66}(3-4), 333--342.

\leavevmode\vadjust pre{\hypertarget{ref-Becker2022OptPre}{}}%
Becker, D. J., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L. M.,
Chen, B., Cohen, L. E., Dallas, T. A., Eskew, E. A., Fagre, A. C.,
Farrell, M. J., Guth, S., Han, B. A., Simmons, N. B., Stock, M.,
Teeling, E. C., \& Carlson, C. J. (2022). Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe}. \url{https://doi.org/10.1016/S2666-5247(21)00245-7}

\leavevmode\vadjust pre{\hypertarget{ref-Bezanson2017JulFre}{}}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020MljJul}{}}%
Blaom, A. D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D., \&
Vollmer, S. J. (2020). MLJ: A Julia package for composable machine
learning. \emph{Journal of Open Source Software}, \emph{5}(55), 2704.
\url{https://doi.org/10.21105/joss.02704}

\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020FleMod}{}}%
Blaom, A. D., \& Vollmer, S. J. (2020). \emph{Flexible model composition
in machine learning and its implementation in MLJ}.
\url{http://arxiv.org/abs/2012.15505}

\leavevmode\vadjust pre{\hypertarget{ref-Boughorbel2017OptCla}{}}%
Boughorbel, S., Jarray, F., \& El-Anbari, M. (2017). Optimal classifier
for imbalanced data using Matthews Correlation Coefficient metric.
\emph{PloS One}, \emph{12}(6), e0177678.
\url{https://doi.org/10.1371/journal.pone.0177678}

\leavevmode\vadjust pre{\hypertarget{ref-Branco2015SurPre}{}}%
Branco, P., Torgo, L., \& Ribeiro, R. (2015). \emph{A Survey of
Predictive Modelling under Imbalanced Distributions}.
\url{http://arxiv.org/abs/1505.01658}

\leavevmode\vadjust pre{\hypertarget{ref-Canard2014EmpEva}{}}%
Canard, E. F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D., \&
Gravel, D. (2014). Empirical evaluation of neutral interactions in
host-parasite networks. \emph{The American Naturalist}, \emph{183}(4),
468--479. \url{https://doi.org/10.1086/675363}

\leavevmode\vadjust pre{\hypertarget{ref-Caron2022AddElt}{}}%
Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
Addressing the Eltonian shortfall with trait-based interaction models.
\emph{Ecology Letters}, \emph{25}(4), 889--899.
\url{https://doi.org/10.1111/ele.13966}

\leavevmode\vadjust pre{\hypertarget{ref-Chicco2020AdvMat}{}}%
Chicco, D., \& Jurman, G. (2020). The advantages of the Matthews
correlation coefficient (MCC) over F1 score and accuracy in binary
classification evaluation. \emph{BMC Genomics}, \emph{21}(1), 6.
\url{https://doi.org/10.1186/s12864-019-6413-7}

\leavevmode\vadjust pre{\hypertarget{ref-Chicco2021MatCor}{}}%
Chicco, D., Tötsch, N., \& Jurman, G. (2021). The Matthews correlation
coefficient (MCC) is more reliable than balanced accuracy, bookmaker
informedness, and markedness in two-class confusion matrix evaluation.
\emph{BioData Mining}, \emph{14}, 13.
\url{https://doi.org/10.1186/s13040-021-00244-z}

\leavevmode\vadjust pre{\hypertarget{ref-deAguiar2019RevBia}{}}%
de Aguiar, M. A. M., Newman, E. A., Pires, M. M., Yeakel, J. D.,
Boettiger, C., Burkle, L. A., Gravel, D., Guimarães, P. R., O'Donnell,
J. L., Poisot, T., Fortin, M.-J., \& Hembry, D. H. (2019). Revealing
biases in the sampling of ecological interaction networks. \emph{PeerJ},
\emph{7}, e7566. \url{https://doi.org/10.7717/peerj.7566}

\leavevmode\vadjust pre{\hypertarget{ref-Delgado2019WhyCoh}{}}%
Delgado, R., \& Tibau, X.-A. (2019). Why Cohen's Kappa should be avoided
as performance measure in classification. \emph{PloS One}, \emph{14}(9),
e0222916. \url{https://doi.org/10.1371/journal.pone.0222916}

\leavevmode\vadjust pre{\hypertarget{ref-Delmas2018AnaEco}{}}%
Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
Fortin, M.-J., Gravel, D., Guimarães, P. R., Hembry, D. H., Newman, E.
A., Olesen, J. M., Pires, M. M., Yeakel, J. D., \& Poisot, T. (2018).
Analysing ecological networks of species interactions. \emph{Biological
Reviews}, 112540. \url{https://doi.org/10.1111/brv.12433}

\leavevmode\vadjust pre{\hypertarget{ref-Desjardins-Proulx2017EcoInt}{}}%
Desjardins-Proulx, P., Laigle, I., Poisot, T., \& Gravel, D. (2017).
Ecological interactions and the Netflix problem. \emph{PeerJ},
\emph{5}(e3644). \url{https://doi.org/10.7717/peerj.3644}

\leavevmode\vadjust pre{\hypertarget{ref-Ferri2009ExpCom}{}}%
Ferri, C., Hernández-Orallo, J., \& Modroiu, R. (2009). An experimental
comparison of performance measures for classification. \emph{Pattern
Recognition Letters}, \emph{30}(1), 27--38.
\url{https://doi.org/10.1016/j.patrec.2008.08.010}

\leavevmode\vadjust pre{\hypertarget{ref-He2013ImbLea}{}}%
He, H., \& Ma, Y. (Eds.). (2013). \emph{Imbalanced Learning:
Foundations, Algorithms, and Applications} (1st edition). Wiley-IEEE
Press.

\leavevmode\vadjust pre{\hypertarget{ref-Hong2016DeaCla}{}}%
Hong, C., Ghosh, R., \& Srinivasan, S. (2016). \emph{Dealing with Class
Imbalance using Thresholding}.
\url{https://doi.org/10.48550/arXiv.1607.02705}

\leavevmode\vadjust pre{\hypertarget{ref-Inman2021ComSam}{}}%
Inman, R., Franklin, J., Esque, T., \& Nussear, K. (2021). Comparing
sample bias correction methods for species distribution modeling using
virtual species. \emph{Ecosphere}, \emph{12}(3), e03422.
\url{https://doi.org/10.1002/ecs2.3422}

\leavevmode\vadjust pre{\hypertarget{ref-Iturbide2015FraSpe}{}}%
Iturbide, M., Bedia, J., Herrera, S., del Hierro, O., Pinto, M., \&
Gutiérrez, J. M. (2015). A framework for species distribution modelling
with improved pseudo-absence generation. \emph{Ecological Modelling},
\emph{312}, 166--174.
\url{https://doi.org/10.1016/j.ecolmodel.2015.05.018}

\leavevmode\vadjust pre{\hypertarget{ref-Japkowicz2013AssMet}{}}%
Japkowicz, N. (2013). Assessment Metrics for Imbalanced Learning. In
\emph{Imbalanced Learning} (pp. 187--206). John Wiley \& Sons, Ltd.
\url{https://doi.org/10.1002/9781118646106.ch8}

\leavevmode\vadjust pre{\hypertarget{ref-Jeni2013FacImb}{}}%
Jeni, L. A., Cohn, J. F., \& De La Torre, F. (2013). Facing Imbalanced
Data--Recommendations for the Use of Performance Metrics. \emph{2013
Humaine Association Conference on Affective Computing and Intelligent
Interaction}, 245--251. \url{https://doi.org/10.1109/ACII.2013.47}

\leavevmode\vadjust pre{\hypertarget{ref-Jordano2016ChaEco}{}}%
Jordano, P. (2016a). Chasing Ecological Interactions. \emph{PLOS Biol},
\emph{14}(9), e1002559.
\url{https://doi.org/10.1371/journal.pbio.1002559}

\leavevmode\vadjust pre{\hypertarget{ref-Jordano2016SamNet}{}}%
Jordano, P. (2016b). Sampling networks of ecological interactions.
\emph{Functional Ecology}. \url{https://doi.org/10.1111/1365-2435.12763}

\leavevmode\vadjust pre{\hypertarget{ref-Landis1977MeaObs}{}}%
Landis, J. R., \& Koch, G. G. (1977). The Measurement of Observer
Agreement for Categorical Data. \emph{Biometrics}, \emph{33}(1),
159--174. \url{https://doi.org/10.2307/2529310}

\leavevmode\vadjust pre{\hypertarget{ref-MacDonald2020RevLin}{}}%
MacDonald, A. A. M., Banville, F., \& Poisot, T. (2020). Revisiting the
Links-Species Scaling Relationship in Food Webs. \emph{Patterns},
\emph{1}(0). \url{https://doi.org/10.1016/j.patter.2020.100079}

\leavevmode\vadjust pre{\hypertarget{ref-Martinez1992ConCon}{}}%
Martinez, N. D. (1992). Constant Connectance in Community Food Webs.
\emph{The American Naturalist}, \emph{139}(6), 1208--1218.
\url{https://www.jstor.org/stable/2462337}

\leavevmode\vadjust pre{\hypertarget{ref-McLeod2021SamAsy}{}}%
McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
asymptotic network properties of spatial multi-trophic networks.
\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}

\leavevmode\vadjust pre{\hypertarget{ref-Olesen2011MisFor}{}}%
Olesen, J. M., Bascompte, J., Dupont, Y. L., Elberling, H., Rasmussen,
C., \& Jordano, P. (2011). Missing and forbidden links in mutualistic
networks. \emph{Proc. R. Soc. B}, \emph{278}(1706), 725--732.
\url{https://doi.org/10.1098/rspb.2010.1371}

\leavevmode\vadjust pre{\hypertarget{ref-Olito2015SpeTra}{}}%
Olito, C., \& Fox, J. W. (2015). Species traits and abundances predict
metrics of plant--pollinator network structure, but not pairwise
interactions. \emph{Oikos}, \emph{124}, 428--436.

\leavevmode\vadjust pre{\hypertarget{ref-Pichler2020MacLea}{}}%
Pichler, M., Boreux, V., Klein, A., Schleuning, M., \& Hartig, F.
(2020). Machine learning algorithms to infer trait‐matching and predict
species interactions in ecological networks. \emph{Methods in Ecology
and Evolution}, \emph{11}(2), 281--293.
\url{https://doi.org/10.1111/2041-210X.13329}

\leavevmode\vadjust pre{\hypertarget{ref-Pilosof2017MulNat}{}}%
Pilosof, S., Porter, M. A., Pascual, M., \& Kéfi, S. (2017). The
multilayer nature of ecological networks. \emph{Nature Ecology \&
Evolution}, \emph{1}, 0101.
\url{https://doi.org/10.1038/s41559-017-0101}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2021GloKno}{}}%
Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
MacDonald, A., Mercier, B., Violet, C., \& Vissault, S. (2021). Global
knowledge gaps in species interaction networks data. \emph{Journal of
Biogeography}, jbi.14127. \url{https://doi.org/10.1111/jbi.14127}

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2021ImpMam}{}}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J.
(2021). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\vadjust pre{\hypertarget{ref-Saito2015PrePlo}{}}%
Saito, T., \& Rehmsmeier, M. (2015). The Precision-Recall Plot Is More
Informative than the ROC Plot When Evaluating Binary Classifiers on
Imbalanced Datasets. \emph{PLOS ONE}, \emph{10}(3), e0118432.
\url{https://doi.org/10.1371/journal.pone.0118432}

\leavevmode\vadjust pre{\hypertarget{ref-Schisterman2005OptCut}{}}%
Schisterman, E. F., Perkins, N. J., Liu, A., \& Bondell, H. (2005).
Optimal Cut-point and Its Corresponding Youden Index to Discriminate
Individuals Using Pooled Blood Samples. \emph{Epidemiology},
\emph{16}(1), 73--81.
\url{https://doi.org/10.1097/01.ede.0000147512.81966.ba}

\leavevmode\vadjust pre{\hypertarget{ref-Somodi2017PreDep}{}}%
Somodi, I., Lepesi, N., \& Botta‐Dukát, Z. (2017). Prevalence dependence
in model goodness measures with special emphasis on true skill
statistics. \emph{Ecology and Evolution}, \emph{7}(3), 863--872.
\url{https://doi.org/10.1002/ece3.2654}

\leavevmode\vadjust pre{\hypertarget{ref-Steen2021SpaThi}{}}%
Steen, V. A., Tingley, M. W., Paton, P. W. C., \& Elphick, C. S. (2021).
Spatial thinning and class balancing: Key choices lead to variation in
the performance of species distribution models with citizen science
data. \emph{Methods in Ecology and Evolution}, \emph{12}(2), 216--226.
\url{https://doi.org/10.1111/2041-210X.13525}

\leavevmode\vadjust pre{\hypertarget{ref-Strona2017ForPer}{}}%
Strona, G., \& Veech, J. A. (2017). Forbidden versus permitted
interactions: Disentangling processes from patterns in ecological
network analysis. \emph{Ecology and Evolution}, \emph{7}(14),
5476--5481. \url{https://doi.org/10.1002/ece3.3102}

\leavevmode\vadjust pre{\hypertarget{ref-Strydom2021RoaPre}{}}%
Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
towards predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, \emph{376}(1837), 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}

\leavevmode\vadjust pre{\hypertarget{ref-Valdovinos2019MutNet}{}}%
Valdovinos, F. S. (2019). Mutualistic networks: Moving closer to a
predictive theory. \emph{Ecology Letters}, \emph{0}(0).
\url{https://doi.org/10.1111/ele.13279}

\leavevmode\vadjust pre{\hypertarget{ref-Weitz2005CoeArm}{}}%
Weitz, J. S., Hartman, H., \& Levin, S. A. (2005). Coevolutionary arms
races between bacteria and bacteriophage. \emph{Proceedings of the
National Academy of Sciences of the United States of America},
\emph{102}(27), 9535--9540.
\url{https://doi.org/10.1073/pnas.0504062102}

\leavevmode\vadjust pre{\hypertarget{ref-Whalen2021NavPit}{}}%
Whalen, S., Schreiber, J., Noble, W. S., \& Pollard, K. S. (2021).
Navigating the pitfalls of applying machine learning in genomics.
\emph{Nature Reviews Genetics}, 1--13.
\url{https://doi.org/10.1038/s41576-021-00434-9}

\leavevmode\vadjust pre{\hypertarget{ref-Wood2015EffSpa}{}}%
Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
\url{https://doi.org/10.1002/ece3.1640}

\leavevmode\vadjust pre{\hypertarget{ref-Xie2017ComCom}{}}%
Xie, J.-R., Zhang, P., Zhang, H.-F., \& Wang, B.-H. (2017). Completeness
of Community Structure in Networks. \emph{Scientific Reports},
\emph{7}(1), 5269. \url{https://doi.org/10.1038/s41598-017-05585-6}

\leavevmode\vadjust pre{\hypertarget{ref-Youden1950IndRat}{}}%
Youden, W. J. (1950). Index for rating diagnostic tests. \emph{Cancer},
\emph{3}(1), 32--35.
\url{https://doi.org/10.1002/1097-0142(1950)3:1\%3C32::AID-CNCR2820030106\%3E3.0.CO;2-3}

\end{CSLReferences}

\end{document}